One-armed spirals in locally isothermal, radially structured self-gravitating discs

### 2015-01-13 paper/discussion.tex
almost done with first revision

%to discuss - fixed central star
%application to ppd
%do smooth disc results carry over to structured disks?

\section{Discussion}\label{discussions} 
{\bf
  We discuss below several issues related to self-gravitating
  discs in the context of our numerical simulations. However, it is
  important to keep in mind that, strictly speaking, 
  the growth of the one-armed spiral in our simulations is
  \emph{not} a gravitational instability in the sense that
  destabilisation is through the background torque associated with a
  forced temperature gradient, and not by self-gravitational torques.     
}

@@

  \subsection{Role of Lindblad and co-rotation torques}
  One effect of self-gravity is to allow the one-armed spiral,
  confined to $R\sim R_0$, to act as an external potential for the
  exterior disc in $R\gtrsim R_0$. This is 
  analogous to disc-satellite interaction 
  \citep{goldreich79}, where the embedded satellite exerts a torque on
  the disc at Lindblad and co-rotation resonances.

@@

  We confirmed this with additional FARGO simulations that exclude the
  co-rotation and outer Lindblad resonances (OLR) by reducing radial domain
  size, setting $R_\mathrm{max} = 3 R_0$. This model still developed
  the one-armed spiral. 
}

{\bf
  \subsection{Applicability to circumstellar discs}

@@

  A locally isothermal equation of state represents the limit
  of infinitely short cooling (and heating) timescales, so the 
  disc temperature instantly returns to its initial value when 
  perturbed. The background torque is generally non-zero if the 
  resulting temperature profile has a non-zero radial gradient.

@@

  % There
  % is no real way to drive a long-lived quasi-steady, non-fragmenting,
  % self-gravitating state because this is only possible if the system
  % can remain quasi-stable though balancing heating and cooling (see,
  % for example, Lodato & Rice 2004).

@@

  In order to confine an $m=1$ mode between two $Q$-barriers, we
  should have $Q^2(1-\nu^2)=1$ at two radii. Assuming 
  Keplerian rotation and a slow pattern speed $\Omega_p\ll\Omega$,
  this amounts to 
  \begin{align}\label{qb_cond}
    \left(\frac{R_{Qb}}{R_c}\right)^{-3/2} = 2Q^2(R_{Qb}). 
  \end{align}
  Then two $Q$-barriers may exist when the $Q^2$ profile
  rises more rapidly (decays more slowly) than $R^{-3/2}$
  for decreasing (increasing) $R$.

@@

%  There is some attempt to do so when discussing dead zones, but I
%  doubt that the conditions there are actually suitable as such a
%  region would either be stable, or cool very slowly and be
%  susceptible to the growth of long-lived, high-m-mode spiral
%  structure.

@@

% The simulations themselves consider an isothermal or locally
% isothermal equation of state only.  Gammie (2001) and other work
% since have shown that the evolution of self-gravitating discs
% depends both on the stability criterion and on the balance between
% heating and cooling.  If the cooling rate is fast, it is likely to
% fragment, otherwise it will settle into a quasi-steady,
% self-gravitating state.

% When considering isothermal simulations, the likely evolution is one
% of two things.  It's either initially so unstable that the disc
% quickly fragments, or it's initially stable enough that any initial
% perturbations simply drives the system towards being even more
% stable and the instability quickly weakens through rapid mass
% transfer (see the work by Laughlin & Bodenheimer 1994/1996).  There
% is no real way to drive a long-lived quasi-steady, non-fragmenting,
% self-gravitating state because this is only possible if the system
% can remain quasi-stable though balancing heating and cooling (see,
% for example, Lodato & Rice 2004).

% In fact, in the conclusions it is suggested that the conditions used
% in the simulation would require a cooling < 0.1 \Omega^{-1} which,
% based on the work of Gammie 2001, would quickly lead to
% fragmentation.  The reason it doesn't in the work considered here is
% that the isothermal assumption assumes that there is some
% equivalently fast heating mechanism but, in the work here, this is
% not provided by the instability itself.  If it were, the simulations
% would become very quickly unstable.  Also, if one looks at the work
% of Clarke (2009), Rafikov (2005), and Rice & Armitage (2009) the
% only region of a realistic disc where such rapid cooling is possible
% is in the outer disc, but such regions are either very
% gravitationally unstable and fragment, or quickly stablised by an
% external heating source.  I guess it is possible that there could be
% a scenario where this exactly produces the conditions assumed here,
% but that would seem rather fine-tuned and it would be more likely
% that an external heating source would simply stabilise the disc
% (unless the disc were extremely massive and could remain unstable
% even at reasonably high disc temperatures).

%Furthermore, this
%non-local interaction may be important for the long-term evolution of
%the outer disc, since it gains angular momentum from the

%also transport ang mom outwards

% Having confirmed the $m=1$ spiral instability through numerical
% simulations, we now use results from the simulations combined with
% linear theory to construct an explanation for their growth. For
% simplicity, we return to 2D discs.

% The dominant disturbance in our simulations is an one-armed, $m=1$
% trailing spiral, and its growth is associated with the imposed temperature
% gradient. As discussed in \S\ref{global_cons}, in a locally isothermal disc this imposed
% temperature gradient generally results in a torque being exerted on
% linear perturbations. An instability is therefore possible if the sign
% of this torque is the same as that of the angular momentum associated
% with the linear perturbation.

%associated angular mometum density is the same as that
%of this
% According to previous studies \citep{adams89}, our disc models
% should not support $m=1$ linear spiral instabilities. This is because
% our discs are not sufficiently massive ($M_d\lesssim 0.1M_*$) and, more
% importantly, we have surpressed the motion of the central star. The 
% latter can induce $m=1$ instabilities in massive discs  \citep{shu90}.

%the latter assumes  k=pi\Sigma G /c_s^2 is always the most favoured

% The temperature gradient affects the growth rate through the
% explicit depdence of the background torque on $q$. The growth 
% rate is also dependent on the underlying disturbance being
% considered (i.e. through $k$ and $\bm{\xi}$ in
% Eq. \ref{theoretical_rate0}). % Of course, the temperature gradient
% % plays no role on local modes. However, other disc parameters such as
% % $c_s^2$ may affect the mode that develops. 
% The dependence on $h$ in 
% Eq. \ref{theoretical_rate} assumes $k=\pi\Sigma G/c_s^2$ is always the
% most favourable wavenumber for low-frequency modes.

%By 
%definition, the global temperature gradient does not affect modes in
%local theory. 
%however, changing $h$ may also modify the mode properties (since
%temperature appears in local theory) in addition
%to the magnitude of the temperature gradient. Thus, we expect the depedence of
%the background torque on $q$ to be better described by
%Eq. \{theoretical_rate}

%by definition, q does not affect local neutral modes
%Of course, we assumed a neutral mode to obtain expressions
%for $\jlin$ and $T_\mathrm{BG}$ in the first place, 
%Recall that in local theory non-axisymmetric modes are stable for real
%$k$  
%Eq. \ref{theoretical_rate} should be interpreted
%as an order-of-magnitude estimate of the perturbative effect of a
%temperature gradient on local, low-frequency $m=1$ modes.

%neutral modes are not actually neutral)
%gamma << omega_p << omega 
%so
%Eq. \ref{theoretical_rate} is only an order-of-magnitude
%constraint. Nevertheless, the expected growth rate is in reasonable
%agreement with numerical results.

%also total torque is integral -> may get cancelations

>>>>

\section{Discussion}\label{discussions} 
{\bf
  We discuss below several issues related to self-gravitating
  discs in the context of our numerical simulations. However, it is
  important to keep in mind that 
  the growth of the one-armed spiral in our simulations is
  \emph{not} a gravitational instability in the sense that
  destabilisation is through the background torque associated with a
  forced temperature gradient, and not by self-gravitational torques\footnote{In fact, additional simulations with $Q_\mathrm{out}=4$ still
  develops the one-armed spiral, but with a smaller growth rate.}. 
}

@@

  \subsection{Role of Lindblad and co-rotation torques}
  One effect of self-gravity is to allow the one-armed spiral,
  confined to $R\sim R_0$ in our models, to act as an external potential for the
  exterior disc in $R>R_0$. This is 
  analogous to disc-satellite interaction 
  \citep{goldreich79}, where the embedded satellite exerts a torque on
  the disc at Lindblad and co-rotation resonances.

@@

  We confirmed this with additional FARGO simulations that exclude the
  co-rotation and outer Lindblad resonances (OLR) by reducing radial domain
  size to $R_\mathrm{max} = 3 R_0$, which still developed 
  the one-armed spiral. 
  %instability due to local exchange of angular momentum with the
  %background disc 
  %CR and LR do not play a role - extended outer disc not essential
  %(but used here to avoid potential boundary effects) cannot exclude
  %a priori 
}

{\bf
  \subsection{Applicability to protoplanetary discs}

@@

  A locally isothermal equation of state represents the ideal limit
  of infinitely short cooling and heating timescales, so the 
  disc temperature instantly returns to its initial value when 
  perturbed. The background torque is generally non-zero if the 
  resulting temperature profile has a non-zero radial gradient.

@@

@@

  In order to confine an $m=1$ mode between two $Q$-barriers, we
  should have $Q^2(1-\nu^2)=1$ at two radii. Assuming 
  Keplerian rotation and a slow pattern speed $\Omega_p\ll\Omega$,
  this amounts to 
  \begin{align}\label{qb_cond}
    \left(\frac{R_{Qb}}{R_c}\right)^{-3/2} = 2Q^2(R_{Qb}). 
  \end{align}
  Then two $Q$-barriers may exist when the $Q^2$ profile
  rises more rapidly (decays more slowly) than $R^{-3/2}$
  for decreasing (increasing) $R$. Note that 
  Eq. \ref{qb_cond} does not necessarily require strong self-gravity
  if $R_c$ is large.

@@

@@

### 2015-01-13 paper/model.tex
almost done with first revision

\section{Governing equations}\label{model} 
We consider an inviscid fluid disc of mass $M_d$ 
orbiting a central star of mass $M_*$. We will mainly examine 2D
(or razor-thin) discs in favour of numerical 
resolution, but have also carried out some 3D disc
simulations. Hence, for generality we describe the system in
3D, using both cylindrical $(R,\phi,z)$ and spherical polar
coordinates $(r,\theta,\phi)$ centred on the star. 
The governing fluid equations in 3D are  
\begin{align}
  &\frac{\p\rho}{\p t} + \nabla\cdot\left(\rho\bm{v}\right) =
  0,\label{cont_eq}\\
  &\frac{\p\bm{v}}{\p t} + \bm{v}\cdot\nabla\bv = -\frac{1}{\rho}\nabla
  p -\nabla \Phi_\mathrm{tot}\label{mom_eq},\\ 
  & \nabla^2\Phi_d = 4 \pi G \rho \label{poisson}, 
\end{align}
where $\rho$ is the mass density, $\bm{v}$ is the 
fluid velocity (the angular velocity being $\Omega \equiv v_\phi/R$), 
$p$ is the pressure and the total potential
$\Phi_\mathrm{tot} = \Phi_* + \Phi_d$ consists of that from the
central star, 
\begin{align}
  \Phi_*(r) = -\frac{GM_*}{r}, 
\end{align}
where $G$ is the gravitational constant,  and the disc potential
$\Phi_d$. We impose a locally isothermal equation of state 
\begin{align}
  p = c_s^2(R)\rho,
\end{align}
where the sound-speed $c_s$ is given by 
\begin{align}\label{sound-speed}
  % c_s(R) = hR_0\Omega_k(R_0)\left(\frac{R}{R_0}\right)^{-q/2}, 
  c_s(R) = c_{s0}\left(\frac{R}{R_0}\right)^{-q/2}
\end{align}
where $c_{s0} = h R_0\Omega_k(R_0)$ and 
$h$ is the disc aspect-ratio at the reference radius $R=R_0$, 
$\Omega_k=\sqrt{GM_*/R^3}$ is the midplane Keplerian frequency, and
$q$ is the imposed temperature gradient (since, for an ideal gas the
temperature is proportional to $c_s^2$). The vertical disc scale-height is
defined by $H=c_s/\Omega_k$. Thus, a strictly isothermal disc with
$q=0$ has $H\propto R^{3/2}$, and $q=1$ corresponds to a 
 disc with constant $H/R$.

@@

However, if the linear disturbance under consideration \emph{is} a
low-frequency mode, then it has negative angular
momentum. If it is trailing and $dc_s^2/dR<0$, as is typical
in astrophysical discs, then $T_\mathrm{BG}<0$, and
the background disc applies a negative torque on the disturbance,
which further decreases its angular momentum. This suggests the {\bf
  mode} 
amplitude will grow.

@@

  There are two components to the destabilising mechanism.  First
  the disc should support low-frequency modes, which is then
  destabilised by an imposed temperature gradient. The latter is
  straight forward to implement by adopting a locally isothermal
  equation of state as described in \S\ref{model}.  For the former, we
  consider discs with a  
  radially-structured Toomre $Q$ profile. We use local theory to show that such discs support
  low-frequency trapped one-armed ($m=1$) modes. 
  This is convenient because Eq. \ref{theoretical_rate} indicates that
  modes with small $m$ are
  more favourable for destabilisation by an imposed temperature
  gradient.  
}

@@

% poisson kernal integration, softening 
% where used?
% BC
\subsubsection{ZEUS-MP}
ZEUS-MP  is a general-purpose finite difference
code \citep{hayes06}. We use the code in 3D spherical geometry, covering
$r\in[r_\mathrm{min},r_\mathrm{max}]$, $\theta\in[\theta_\mathrm{min},\pi/2]$,
$\phi\in[0,2\pi]$. The vertical domain is chosen to cover $n_H$
scale-heights at $R=R_0$, i.e. $\tan{(\pi/2 - \theta_\mathrm{min})}/h=n_H$. 
The grid is logarithmically spaced in radius and uniformly spaced in the angular
coordinates. We assume symmetry across the midplane, and
apply reflective boundary conditions at radial boundaries and the
upper disc boundary.

@@

%used in lin 2012 
\subsubsection{PLUTO} 
PLUTO is a general-purpose Godunov code \citep{mignone07}. The grid
setup is the same as that adopted in ZEUS-MP above. We configure the
code similarly to that used in \cite{lin14}: piece-wise linear
reconstruction, a Roe solver and second order Runge-Kutta time
integration. We also enable the FARGO algorithm for azimuthal
transport.

@@

%used in conference proceeding

@@

The normalised surface density with azimuthal wavenumber $m$ is
\begin{align}
  \Delta\Sigma_m = \frac{2}{\Sigma_{00}} \real\left[\Sigma_m \exp{(\ii
      m\phi)}\right]
\end{align}
where $\Sigma_{00} = \Sigma_0(t=0)$. The time evolution of the
$m^\mathrm{th}$ mode can be characterized by 
assuming $\Sigma_m\propto\exp{(-\ii \sigma t)}$ as in linear
theory. 
%where
%$\sigma = \omega + \ii\gamma$ is a complex frequency, $\omega$ being 
%the real frequency and $\gamma$ is the growth rate. 
The total non-axisymmetric surface density is 
\begin{align}
  \Delta\Sigma = \frac{\Sigma - \Sigma_0}{\Sigma_0}. 
\end{align}

\subsubsection{Angular momentum decomposition}
The total disc angular momentum is
\begin{align}
  J &= \int_{R_\mathrm{min}}^{R_\mathrm{max}}\int_0^{2\pi} \Sigma Rv_\phi RdRd\phi \notag\\
  &= 2\pi\int_{R_\mathrm{min}}^{R_\mathrm{max}} R\Sigma_0 v_{\phi0} R dR \notag\\ 
  &\phantom{=}+
  \sum_{m=1}^\infty2\pi\int_{R_\mathrm{min}}^{R_\mathrm{max}} 2R\real\left[\Sigma_m v_{\phi
      m}^*\right] RdR 
  %&= \sum_{m=0}^\infty \int _{R_\mathrm{min}}^{R_\mathrm{max}}  j_m(R)
  %dR 
  = \sum_{m=0}^\infty J_m. 
\end{align}
We will refer to $J_m$ as the
$m^\mathrm{th}$ component of the total angular momentum, and use it to
monitor numerical angular momentum conservation in the simulations. 
It is important to distinguish this empirical definition from the
angular momentum of linear perturbations given in \S\ref{wkb}, which
is defined through a conservation law.

%and we idenfity $J_m$ as the  
%angular momentum associated with disc structure with
%$m^\mathrm{th}$-fold symmetry, and $j_m$ the associated angular
%momentum per unit length. We will use $J_m$ to monitor angular momentum 
%conservation for the disc as a whole.

@@

% \begin{align}
%   T_\mathrm{BG} &= -\frac{m}{2}\frac{dc_s^2}{dR}\imag\left[\delta\Sigma_m \left(\frac{\delta
%         v_{Rm}}{-\ii\sbar}\right)^*\right] \notag\\
%   & = \frac{m}{2}\frac{dc_s^2}{dR}\imag\left[\ii \delta\Sigma_m \left(\frac{\delta
%         v_{Rm}}{\sbar}\right)^*\right]
% \end{align}
% where we replaced the radial Lagrangian displacement by the Eulerian
% radial velocity perturbation in Eq. \ref{baroclinic_torque}. 
% The linearised momentum equations give
% \begin{align}
%   D\delta v_{Rm} =& 
%   \ii\sbar\left[c_s^2\frac{d}{dR}\left(\frac{\delta\Sigma_m}{\Sigma}\right)
%     + \frac{d}{dR}\delta\Phi_m\right] \notag\\ 
%   &- \frac{2\ii
%     m\Omega}{R}\left(c_s^2\frac{\delta\Sigma_m}{\Sigma} +
%     \delta\Phi_m\right),
% \end{align}
% where $D\equiv \kappa^2 - \sbar^2$. 
% We now invoke local theory by setting $d/dR \to \ii k$ where $k$ is
% real, and using the local solution to the Poisson equation
% \begin{align}
%   \delta\Phi_m = -\frac{2\pi G}{|k|}\delta\Sigma_m 
% \end{align}
% \citep{shu91}. Then the radial velocity perturbation becomes
% \begin{align}
%   \delta v_{Rm} =
%   \frac{1}{D}\frac{\delta\Sigma_m}{\Sigma}\left(\frac{2\pi G
%       \Sigma}{|k|} - c_s^2\right)\left(k\sbar + \frac{2\ii
%       m\Omega}{R}\right),
% \end{align}
% so that 
% \begin{align}
%   \ii\delta\Sigma_m \left(\frac{\delta v_{Rm}}{\sbar}\right)^* =
%   \frac{1}{D^*}\frac{|\delta\Sigma_m|^2}{\Sigma}\left(\frac{2\pi G
%       \Sigma}{|k|} - c_s^2\right)\left(\ii k  + \frac{2
%       m\Omega}{R\sbar^*}\right). 
% \end{align}
% For a neutral mode, $\sbar$, and therefore $D$ is real. Then, after
% taking the imaginary part of the above expression, the torque
% density is 
% \begin{align}\label{tbg_explicit}
%   T_\mathrm{BG} &=
%   \frac{m}{2}\frac{dc_s^2}{dR}\left[\frac{k}{D}\frac{|\delta\Sigma_m|^2}{\Sigma}\left(\frac{2\pi G
%       \Sigma}{|k|} - c_s^2\right)\right].
% % &\simeq \frac{m}{2}\frac{dc_s^2}{dR}\left[\frac{k}{\left(\kappa^2 -
% %         m^2\Omega^2\right)}\frac{|\delta\Sigma_m|^2}{\Sigma}\left(\frac{2\pi
% %       G \Sigma}{|k|} - c_s^2\right)\right]. 
% \end{align}
% We now use the local dispersion relation,
% Eq. \ref{dispersion}, to make the replacement $D = 2\pi G \Sigma |k| -
% k^2c_s^2$ in the above expression. This gives
% \begin{align}\label{tbg_simple}
%   T_\mathrm{BG} =
%   \frac{m}{2}\frac{dc_s^2}{dR}\frac{1}{k}\frac{|\delta\Sigma_m|^2}{\Sigma}.  
% \end{align}

>>>>

\section{Governing equations}\label{model} 
We consider an inviscid fluid disc of mass $M_d$ 
orbiting a central star of mass $M_*$. We will mainly examine 2D
(or razor-thin) discs in favour of numerical 
resolution, but have also carried out some 3D disc
simulations. Hence, for generality we describe the system in
3D, using both cylindrical $(R,\phi,z)$ and spherical polar
coordinates $(r,\theta,\phi)$ centred on the star. 
The governing fluid equations in 3D are  
\begin{align}
  &\frac{\p\rho}{\p t} + \nabla\cdot\left(\rho\bm{v}\right) =
  0,\label{cont_eq}\\
  &\frac{\p\bm{v}}{\p t} + \bm{v}\cdot\nabla\bv = -\frac{1}{\rho}\nabla
  p -\nabla \Phi_\mathrm{tot}\label{mom_eq},\\ 
  & \nabla^2\Phi_d = 4 \pi G \rho \label{poisson}, 
\end{align}
where $\rho$ is the mass density, $\bm{v}$ is the 
fluid velocity (the angular velocity being $\Omega \equiv v_\phi/R$), 
$p$ is the pressure and the total potential
$\Phi_\mathrm{tot} = \Phi_* + \Phi_d$ consists of that from the
central star, 
\begin{align}
  \Phi_*(r) = -\frac{GM_*}{r}, 
\end{align}
where $G$ is the gravitational constant,  and the disc potential
$\Phi_d$. We impose a locally isothermal equation of state 
\begin{align}
  p = c_s^2(R)\rho,
\end{align}
where the sound-speed $c_s$ is given by 
\begin{align}\label{sound-speed}
  % c_s(R) = hR_0\Omega_k(R_0)\left(\frac{R}{R_0}\right)^{-q/2}, 
  c_s(R) = c_{s0}\left(\frac{R}{R_0}\right)^{-q/2}
\end{align}
where $c_{s0} = h R_0\Omega_k(R_0)$ and 
$h$ is the disc aspect-ratio at the reference radius $R=R_0$, 
$\Omega_k=\sqrt{GM_*/R^3}$ is the midplane Keplerian frequency, and
$q$ is the imposed temperature gradient since, for an ideal gas the
temperature is proportional to $c_s^2$. {\bf For convenience we will
  refer to $c_s^2$ as the disc temperature.} The vertical disc scale-height is
defined by $H=c_s/\Omega_k$. Thus, a strictly isothermal disc with
$q=0$ has $H\propto R^{3/2}$, and $q=1$ corresponds to a 
 disc with constant $H/R$.

@@

However, if the linear disturbance under consideration \emph{is} a
low-frequency mode, then it has negative angular
momentum. If it is trailing and $dc_s^2/dR<0$, as is typical
in astrophysical discs, then $T_\mathrm{BG}<0$ and
the background disc applies a negative torque on the disturbance,
which further decreases its angular momentum. This suggests the {\bf
  mode} 
amplitude will grow.

@@

  There are two parts to the destabilising mechanism: 
  the disc should support low-frequency modes, which is then
  destabilised by an imposed temperature gradient. The latter is
  straight forward to implement by adopting a locally isothermal
  equation of state as described in \S\ref{model}.  For the former, we
  consider discs with a  
  radially-structured Toomre $Q$ profile. We use local theory to show
  that such discs can trap 
  low-frequency one-armed ($m=1$) modes.  
  This is convenient because Eq. \ref{theoretical_rate} indicates that
  modes with small $m$ are
  more favourable for destabilisation. 
}

@@

\subsubsection{ZEUS-MP}
ZEUS-MP  is a general-purpose finite difference
code \citep{hayes06}. We use the code in 3D spherical geometry, covering
$r\in[r_\mathrm{min},r_\mathrm{max}]$, $\theta\in[\theta_\mathrm{min},\pi/2]$,
$\phi\in[0,2\pi]$. The vertical domain is chosen to cover $n_H$
scale-heights at $R=R_0$, i.e. $\tan{(\pi/2 - \theta_\mathrm{min})}/h=n_H$. 
The grid is logarithmically spaced in radius and uniformly spaced in the angular
coordinates. We assume symmetry across the midplane, and
apply reflective boundary conditions at radial boundaries and the
upper disc boundary.

@@

\subsubsection{PLUTO} 
PLUTO is a general-purpose Godunov code \citep{mignone07}. The grid
setup is the same as that adopted in ZEUS-MP above. We configure the
code similarly to that used in \cite{lin14}: piece-wise linear
reconstruction, a Roe solver and second order Runge-Kutta time
integration. We also enable the FARGO algorithm for azimuthal
transport.

@@

@@

The normalised surface density with azimuthal wavenumber $m$ is
\begin{align}
  \Delta\Sigma_m = \frac{2}{\Sigma_{00}} \real\left[\Sigma_m \exp{(\ii
      m\phi)}\right]
\end{align}
where $\Sigma_{00} = \Sigma_0(t=0)$. The time evolution of the
$m^\mathrm{th}$ mode can be characterized by 
assuming $\Sigma_m\propto\exp{(-\ii \sigma t)}$ as in linear
theory.

The total non-axisymmetric surface density is 
\begin{align}
  \Delta\Sigma = \frac{\Sigma - \Sigma_0}{\Sigma_0}. 
\end{align}

\subsubsection{Angular momentum decomposition}
The total disc angular momentum is
\begin{align}
  J &= \int_{R_\mathrm{min}}^{R_\mathrm{max}}\int_0^{2\pi} \Sigma Rv_\phi RdRd\phi \notag\\
  &= 2\pi\int_{R_\mathrm{min}}^{R_\mathrm{max}} R\Sigma_0 v_{\phi0} R dR \notag\\ 
  &\phantom{=}+
  \sum_{m=1}^\infty2\pi\int_{R_\mathrm{min}}^{R_\mathrm{max}} 2R\real\left[\Sigma_m v_{\phi
      m}^*\right] RdR 
  = \sum_{m=0}^\infty J_m. 
\end{align}
We will refer to $J_m$ as the
$m^\mathrm{th}$ component of the total angular momentum, and use it to
monitor numerical angular momentum conservation in the simulations. 
It is important to distinguish this empirical definition from the
angular momentum of linear perturbations given in \S\ref{wkb}, which
is defined through a conservation law.

@@
